Parallel Tractability of Ontology\\ Materialization: Technique and Practice

### 2018-02-01 parallel-tractability-J-v9/evaluation.tex
worked on experimental section, fixed table layout and captions

In the first part of this section, we analyze different kinds of ontologies and datasets
and investigate the usability of the results proposed above.
In the second part, we evaluate the implementation of Algorithm~$\mathsf{A}_{prc}$
and compare it to the reasoning system RDFox.

@@

\textbf{Benchmarks}. In the
Semantic Web community, many benchmarks are proposed to
facilitate the evaluation of ontology-based systems
in a standard and systematic way. We investigate several
popular benchmarks using our results and find that
the ontologies used in some benchmarks have simple
structured TBoxes that can be expressed in RDFS
and belong to $\mathcal{D}_{\textit{\text{dhl}}}$.
These benchmarks include SIB\footnote{https://www.w3.org/wiki/Social$\underline{~}$Network$\underline{~}$Intelligence$\underline{~}$BenchMark}
(\emph{Social Network Intelligence BenchMark}),
BSBM\footnote{http://wifo5-03.informatik.uni-mannheim.de/bizer/berlinsparqlbenchmark/}
(\emph{Berlin SPARQL Benchmark}) and
LODIB\footnote{http://wifo5-03.informatik.uni-mannheim.de/bizer/lodib/}
(\emph{Linked Open Data Integration Benchmark}).
The ontology used in IIMB\footnote{http://islab.di.unimi.it/iimb/}
(\emph{The ISLab Instance Matching Benchmark}) follows the simple-concept restriction.

In the latest version of LUBM\footnote{http://swat.cse.lehigh.edu/projects/lubm/}
(The Lehigh University Benchmark),
there are 48 classes and 32 properties.
Statements about properties, such as inverse property statements,
can be rewritten into datalog rules allowed in $\mathcal{D}_{\textit{\text{dhl}}}$.
Most of the statements about classes can be rewritten into datalog rules that
are allowed in $\mathcal{D}_{\textit{\text{dhl}}}$.
Five axioms have, however, the form $A\sqsubseteq\exists R.B$,
which requires existentially quantified variables in the rule head when rewriting
the axiom into a logic rule: $A(x)\rightarrow\exists y(R(x,y)\wedge B(y))$,
where a new anonymous constant $y$ is introduced.
This kind of rule is not considered when using OWL RL reasoners to handle LUBM \cite{UrbaniKMHB12,WeaverH09}.
On the other hand, in some cases, this kind of rule can also be
eliminated when taking a rewriting approach \cite{GrauHKKMMW13}.
In summary, if the above kind of rule is not considered,
the materialization of a LUBM dataset can be handled by Algorithm~$\mathsf{A}_{opt}^{\psi}$.

\textbf{YAGO}. The knowledge base YAGO\footnote{http://www.mpi-inf.mpg.de/home/}
is constructed from Wikipedia and WordNet and the latest version
YAGO3 \cite{MahdisoltaniBS15} has more than 10 million entities
(e.g., persons, organizations, cities, etc.)
and contains more than 120 million facts about these entities.
In order to balance the expressiveness and computing efficiency,
a YAGO-style language, called YAGO \emph{model}, is proposed based on
a slight extension of RDFS \cite{SuchanekKW08}.
In addition to the expressiveness of RDFS,
YAGO \emph{model} also allows stating the \emph{transitivity} and \emph{acyclicity} of a property.
Making full use of RDFS features cannot lead to parallel tractability \cite{Horst05}.
However, in \cite{SuchanekKW08}, a group of materialization rules
is specified, which is more efficient.
All of these rules are allowed in $\mathcal{D}_{\textit{\text{dhl}}}$.
Thus, we have that a well-constructed YAGO dataset belongs to $\mathcal{D}_{\textit{\text{dhl}}}$.

\textbf{Real Ontologies}. We investigated
$151$ ontologies that cover many domains
like biomedicine, geography, etc.
These ontologies are collected
from the Protege ontology library,\footnote{http://protegewiki.stanford.edu/wiki/Protege$\underline{~}$Ontology$\underline{~}$Library}
Swoogle\footnote{http://swoogle.umbc.edu/}
and Oxford ontology lib.\footnote{http://www.cs.ox.ac.uk/isg/ontologies/lib/}
All ontologies are available online.\footnote{https://github.com/quanzz/PT}
Among these ontologies, $111$ of them belong to $\mathcal{D}_{\textit{\text{dhl}}}$
or $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$, and $21$ DHL ontologies contain
conjunctions and follow the simple-concept restriction.
The remaining ontologies have simple TBoxes, i.e., no conjunction ($A_1\sqcap A_2$)
appears in these ontologies.
We also find two DHL($\circ$) ontologies
that follow the simple-role restriction.

The above investigation indicates that the simple-concept and the simple-role restrictions
are allowed in many real applications of different fields.
From the perspective of developers who work on building their own ontologies,
they can also refer to $\mathcal{D}_{\textit{\text{dhl}}}$ and $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$
to achieve the guarantee of the parallel tractability.

\subsection{Evaluating the Implementation of Algorithm~$\mathsf{A}_{prc}$}

We implemented a prototype system ParallelDHL for DHL$(\circ)$ materialization
based on Algorithm~$\mathsf{A}_{prc}$. In this part, we evaluate ParallelDHL and
compare it to the state-of-the-art reasoning system, RDFox \cite{MotikNPHO14},
that can also handle ontology materialization.

\textbf{Datasets}.
We select eight ontologies from the data sources given in the previous subsection;
four of the eight ontologies belong to $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$ and
the rest of them do not follow the simple-concept or the simple-role
restriction.
The eight ontologies are real ontologies and are applied in different fields.
The basic information of these ontologies
is summarized in Table~2 (we use $\mathcal{D}^-_{\textit{\text{dhl}}(\circ)}$
to denote the complementary set of $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$).

\begin{center}
\begin{tabular}{ccrrc}
\multicolumn{5}{c}{\textbf{Table 2: The Information of the Test Ontologies}}\\
\hline
~~~~~\textbf{Ontology Name}~~~~~&~~~~~\textbf{Field}~~~~~&~~~~~$|\mathcal{T}|+|\mathcal{R}|^a$&~~~~~$|\mathcal{A}|^b$&~~~~~\textbf{Type}$^c$~~~~~\\
\hline

Finance&finance&1,934&6,152&$\mathcal{D}^-_{\textit{\text{dhl}}(\circ)}$\\

Molecule&chemistry&43&0&$\mathcal{D}^-_{\textit{\text{dhl}}(\circ)}$\\

GrossAnatomy&anatomy&2,276&13&$\mathcal{D}^-_{\textit{\text{dhl}}(\circ)}$\\

Skeleton&medicine&815&0&$\mathcal{D}^-_{\textit{\text{dhl}}(\circ)}$\\

@@

ChemistryPrimitive&chemistry&167&0&$\mathcal{D}_{\textit{\text{dhl}}(\circ)}$\\

FacebookOnto&social networking&185&28&$\mathcal{D}_{\textit{\text{dhl}}(\circ)}$\\

Mahabharata&literature&69&2,036&$\mathcal{D}_{\textit{\text{dhl}}(\circ)}$\\

Transportation&traffic&925&511&$\mathcal{D}_{\textit{\text{dhl}}(\circ)}$\\

\hline
\multicolumn{5}{l}{$^a$ \footnotesize$|\mathcal{T}|+|\mathcal{R}|$ denotes the number of axioms occurring in the ontology.}\\

\multicolumn{5}{l}{$^b$ \footnotesize$|\mathcal{A}|$ denotes the number of assertions occurring in the ABox.}\\

\multicolumn{5}{l}{$^c$ \footnotesize$\mathcal{D}_{\textit{\text{dhl}}(\circ)}$ (resp., $\mathcal{D}^-_{\textit{\text{dhl}}(\circ)}$)
means that the ontology belongs to $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$ (resp., $\mathcal{D}^-_{\textit{\text{dhl}}(\circ)}$).}\\
\end{tabular}
\label{tab:info}
\end{center}

For the purpose of comparability in the ABox scale, we consider generating ABox assertions
to a limited number with respect to the TBoxes and RBoxes of the test ontologies.
The generation method is based on the work \cite{Elhaik98} where a benchmark generator is proposed
for evaluating ontology-based systems.
For each test ontology, e.g., the ontology Finance, we generate 5 new ontologies with different ABoxes;
these ontologies are denoted by Finance-$i$ ($i\in\{1,2,3,4,5\}$) where $i$ represents that the number
of assertions in the ABox of Finance-$i$ reaches to $i\times100,000$;
we also use ``Finance series" to denote the five generated ontologies
for the ontology Finance. The other test ontologies are processed similarly.

\textbf{The Experimental Results}.
We run ParallelDHL and RDFox over the above eight ontology series respectively.
The running environment is a DELL server with a
memory of 16 GiB and 4 cores.
For fairness, we set the same number of threads (i.e., 1,2,4,6 and 8 threads respectively)
to ParallelDHL and RDFox in each experiment. The results of reasoning
times\footnote{The experimental results can be found at https://github.com/quanzz/PT.} are
presented in the 16 line graphs (denoted by $lg1$,...,$lg16$) of Figure~\ref{fig:linegraph},
where the abscissa of each line graph records the numbers of ABox assertions,
the ordinate records the values of reasoning time (millisecond per unit),
and each of the five curves in different colors denotes the trend of reasoning time with the
corresponding number of threads allocated (we use line-$k$ to denote the curve
corresponding to $k$ threads).

We further process the collected data and fill the results in Table~3, where each cell corresponds
to a test ontology (see the row label) and a new generated ABox (distinguished by the column labels);
the three values from above to below in each cell are: (1) \emph{the minimal time ratio} - the minimal reasoning time
of ParallelDHL divided by the minimal reasoning time of RDFox;
(2) \emph{the average speedup of ParallelDHL};\footnote{Suppose $T_i$ is the reasoning time with $i$ threads allocated,
the average speedup is $\frac{1}{4}(\frac{T_1}{T_2}+\frac{T_2}{T_4}+\frac{T_4}{T_6}+\frac{T_6}{T_8})$.}
(3) \emph{the average speedup of RDFox}.
The minimal time ratio describes the performance of ParallelDHL by using RDFox as the baseline.
The indicator speedup and its derived indicators are the most common tools used for measuring the capacity of parallelism \cite{MotikNPHO14,KazakovKS14,UrbaniKMHB12}.
Here, we use the average speedup \cite{ichiyoshiK92} to describe the average capacity of parallelism with different
threads allocated.
In the following, we give the detailed analysis based on the contents in Figure~\ref{fig:linegraph} and Table~3.

\textbf{The Analysis of the Experimental Results}.
According to the theoretical results in Section~\ref{sec:ptonto}, the ontologies belonging to $\mathcal{D}^-_{\textit{\text{dhl}}(\circ)}$
may not be parallelly tractable in terms of materialization. In other words,
parallel techniques may not work for improving the efficiency of materialization.
This is shown in the line graphs of Figure~\ref{fig:linegraph}.
We can see that, in the line graphs ($lg1$-$lg8$) of the ontology series
that belong to $\mathcal{D}^-_{\textit{\text{dhl}}(\circ)}$, the four lines, line-2, line-4, line-6 and line-8,
intersect to some degree. This situation in the line graphs for Finance and GrossAnatomy
series is more obvious. For example, in line graph $lg1$,
line-2 stays higher than line-6, but lower than line-4.
The intersection of lines indicates that reasoning time cannot be obviously reduced with more threads allocated.
This is also supported by the results of the average speedups.
From Table~3, we can see that the average speedups of ParallelDHL and RDFox
for ChemistryPrimitive, FacebookOnto, Mahabharata and Transportation series
are more stable compared with the other ones, i.e., they are averagely 1.4
(it means that the reasoning time can be reduced averagely by 1.4 times with two more threads allocated)
for ParallelDHL, and, 1.1 for RDFox. For the four ontology series that belong to $\mathcal{D}^-_{\textit{\text{dhl}}(\circ)}$,
the average speedups have a higher volatility. For example,
Finance and Skeleton series lead to high average speedups that reach up to 3,
while, for Molecule and GrossAnatomy series, the average speedups are even lower than 1.
In summary, from the experiments on the test ontology series,
parallelism leads to a more effective improvement for materializing the ontology series
that belong to $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$ compared with the ones in $\mathcal{D}^-_{\textit{\text{dhl}}(\circ)}$.

By analyzing the experimental results of ParallelDHL and RDFox, we have that
ParallelDHL is a competitive system.
From Table~3, we can find that most of the minimal time ratios are close to 1.
This means that the reasoning times of ParallelDHL are close to that of RDFox in parallel.
In particular, for GrossAnatomy, FacebookOnto and Transportation series, ParallelDHL performs
much better than RDFox; therein, for materializing FacebookOnto-3, ParallelDHL
only costs one fifth of the minimal reasoning time of RDFox.
On the other hand, for most ontology series,
the average speedups of ParallelDHL are overall higher than that of RDFox.
The main difference of ParallelDHL and RDFox lies in that ParallelDHL applies
the optimizations designed for Algorithm~$\mathsf{A}_{opt}$ (see Section~\ref{sec:ptclass}).
The higher average speedups of PrallelDHL also verifies the validity of the optimizations
used in Algorithm~$\mathsf{A}_{opt}$.
One may note that, ParallelDHL is averagely 3 times slower than RDFox when handling Skeleton series.
The main reason is that the computation of \texttt{rch} relations
occupies a large amount of time. We have checked that, the situation of path twisting occurs in Skeleton series.
This makes the the optimization based on \texttt{rch} relations invalid as discussed
in Section~\ref{sec:ptonto}.

>>>>

In the first part of this section, we analyze different kinds of ontologies and datasets
and investigate the usability of the results proposed above.
In the second part, we evaluate the implementation of Algorithm~$\mathsf{A}_{\text{prc}}$
and compare it to the reasoning system RDFox.

@@

\textbf{Benchmarks}. In the
Semantic Web community, many benchmarks are proposed to
facilitate the evaluation of ontology-based systems
in a standard and systematic way. We investigate several
popular benchmarks using our results. For some benchmarks, the ontologies have TBoxes with a simple
structure. These ontologies can be expressed in RDFS
and belong to $\mathcal{D}_{\textit{\text{dhl}}}$.
These benchmarks include SIB\footnote{https://www.w3.org/wiki/Social$\underline{~}$Network$\underline{~}$Intelligence$\underline{~}$BenchMark}
(\emph{Social Network Intelligence BenchMark}),
BSBM\footnote{http://wifo5-03.informatik.uni-mannheim.de/bizer/berlinsparqlbenchmark/}
(\emph{Berlin SPARQL Benchmark}) and
LODIB\footnote{http://wifo5-03.informatik.uni-mannheim.de/bizer/lodib/}
(\emph{Linked Open Data Integration Benchmark}).
The ontology used in IIMB\footnote{http://islab.di.unimi.it/iimb/}
(\emph{The ISLab Instance Matching Benchmark}) satisfies the simple-concept restriction.

In the latest version of LUBM\footnote{http://swat.cse.lehigh.edu/projects/lubm/}
(The Lehigh University Benchmark),
there are 48 classes and 32 properties.
Statements about properties, such as inverse property statements,
can be rewritten into datalog rules allowed in $\mathcal{D}_{\textit{\text{dhl}}}$.
Most of the statements about classes can be rewritten into datalog rules that
are allowed in $\mathcal{D}_{\textit{\text{dhl}}}$.
Five axioms have, however, the form $A\sqsubseteq\exists R.B$,
which requires existentially quantified variables in the rule head when rewriting
the axiom into a logic rule: $A(x)\rightarrow\exists y(R(x,y)\wedge B(y))$,
where a new anonymous constant $y$ is introduced.
This kind of rule is not considered when using OWL RL reasoners to handle LUBM \cite{UrbaniKMHB12,WeaverH09}.
On the other hand, in some cases, this kind of rule can also be
eliminated when taking a rewriting approach \cite{GrauHKKMMW13}.
In summary, if the above kind of rule is not considered,
the materialization of a LUBM dataset can be handled by Algorithm~$\mathsf{A}_{\text{opt}}^{\psi}$.

\textbf{YAGO}. The knowledge base YAGO\footnote{http://www.mpi-inf.mpg.de/home/}
is constructed from Wikipedia and WordNet and the latest version
YAGO3 \cite{MahdisoltaniBS15} has more than 10 million entities
(e.g., persons, organizations, cities, etc.)
and contains more than 120 million facts about these entities.
In order to balance the expressiveness and computing efficiency,
a YAGO-style language, called YAGO \emph{model}, is proposed based on
a slight extension of RDFS \cite{SuchanekKW08}.
In addition to the expressiveness of RDFS,
YAGO \emph{model} also allows for stating the \emph{transitivity} and \emph{acyclicity} of a property.
Making full use of RDFS features cannot lead to parallel tractability \cite{Horst05}.
However, \citet{SuchanekKW08} specify a more efficient set of materialization rules.
All of these rules are allowed in $\mathcal{D}_{\textit{\text{dhl}}}$.
Thus, we have that a well-constructed YAGO dataset belongs to $\mathcal{D}_{\textit{\text{dhl}}}$.

\textbf{Real Ontologies}. We investigated
$151$ ontologies that cover many domains
like biomedicine, geography, etc.
These ontologies are collected
from the Protege ontology library,\footnote{http://protegewiki.stanford.edu/wiki/Protege$\underline{~}$Ontology$\underline{~}$Library}
Swoogle\footnote{http://swoogle.umbc.edu/}
and the Oxford ontology library.\footnote{http://www.cs.ox.ac.uk/isg/ontologies/lib/}
All ontologies are available online.\footnote{https://github.com/quanzz/PT}
Among these ontologies, $111$ belong to $\mathcal{D}_{\textit{\text{dhl}}}$
or $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$; $21$ of the DHL ontologies contain
conjunctions and satisfy the simple-concept restriction.
The remaining ontologies have simple TBoxes, i.e., no conjunction ($A_1\sqcap A_2$)
appears in these ontologies.
We also find two DHL($\circ$) ontologies and both satisfy the simple-role restriction.

The above investigation indicates that the simple-concept and the simple-role restrictions
is naturally satisfied in many real applications of different fields.
From the perspective of developers who work on building their own ontologies,
they can also refer to $\mathcal{D}_{\textit{\text{dhl}}}$ and $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$
to achieve the guarantee of the parallel tractability.

\subsection{Evaluating the Implementation of Algorithm~$\mathsf{A}_{\text{prc}}$}

We implemented a prototype system ParallelDHL for DHL$(\circ)$ materialization
based on Algorithm~$\mathsf{A}_{\text{prc}}$. In this part, we evaluate ParallelDHL and
compare it to the state-of-the-art reasoning system, RDFox \cite{MotikNPHO14},
that can also handle ontology materialization.

\textbf{Datasets}.
We select eight ontologies from the data sources given in the previous subsection;
four of the eight ontologies belong to $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$ and
the rest of them does not satisfy the simple-concept or the simple-role
restriction.
The chosen eight ontologies are real ontologies and are applied in different fields.
The basic information for these ontologies
is summarized in Table~\ref{tab:info}.

\begin{table}[htb]
\centering
\caption{Information about the test ontologies, where
  $|\mathcal{T}|+|\mathcal{R}|$ denotes the number of axioms occurring
  in the ontology, $|\mathcal{A}|$ denotes the number of assertions
  in the ABox and the last column states whether the
  ontology belongs to $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$}
\begin{tabular}{>{\hspace*{5mm}}ccrrc<{\hspace*{5mm}}}
\hline
\textbf{Ontology Name} & \textbf{Field} &
                                          $|\mathcal{T}|+|\mathcal{R}|$
  & $|\mathcal{A}|$ & $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$\\
\hline

Finance&finance&1,934&6,152&no\\

Molecule&chemistry&43&0&no\\

GrossAnatomy&anatomy&2,276&13&no\\

Skeleton&medicine&815&0&no\\

@@

ChemistryPrimitive&chemistry&167&0& yes\\

FacebookOnto&social networking&185&28& yes\\

Mahabharata&literature&69&2,036& yes\\

Transportation&traffic&925&511& yes\\

\hline
\end{tabular}
\label{tab:info}
\end{table}

In order to compare ABoxes of a similar scale, we generate a limited
number of ABox assertions with respect to the TBoxes and RBoxes for
the test ontologies using the generation method proposed by
\citet{Elhaik98}.  For each test ontology, e.g., the Finance ontology,
we generate 5 new ontologies with different ABoxes, denoted
Finance-$i$ ($i\in\{1,2,3,4,5\}$). The index $i$ indicates that the
number of ABox assertions of Finance-$i$ is $i\times100,000$. We also
use the term ``Finance series" to denote the five generated ontologies
for the Finance ontology. The other test ontologies are processed
similarly.

\begin{table}[htb]
\centering
\caption{Analysis of the reasoning results, where the \emph{minimal time
  ratio} is the minimal reasoning time
of ParallelDHL divided by the minimal reasoning time of RDFox and the
\emph{average speedup} is $1/4 (T_1/T_2 + T_2/T_4 + T_4/T_6 + T_6/T_8)$,
 where $T_i$ is the reasoning time with $i$ threads allocated}
{\setlength{\tabcolsep}{6mm}
\begin{tabular}{crrrrr}
\hline
series & 1 & 2 & 3 & 4 & 5\\
\hline
\multicolumn{6}{c}{minimal time ratio}\\%[-1mm]
% \multicolumn{6}{c}{{\footnotesize minimal reasoning time
% of ParallelDHL/minimal reasoning time of RDFox}}\\
\hline
Finance&0.54&1.36&0.94&0.99&1.70\\
Molecule&0.94&0.79&0.96&0.98&1.07\\
GrossAnatomy&0.65&0.48&0.84&0.77&0.83\\
Skeleton&1.11&\textbf{3.08}&\textbf{3.03}&\textbf{3.04}&\textbf{2.75}\\
ChemistryPrimitive&1.29&1.54&1.68&1.14&0.93\\
FacebookOnto&\textbf{0.23}&\textbf{0.22}&\textbf{0.19}&\textbf{0.24}&\textbf{0.22}\\
Mahabharata&1.19&1.19&0.99&0.84&0.89\\
Transportation&0.24&0.52&0.51&0.65&0.49\\
\hline
\multicolumn{6}{c}{average speedup of ParallelDHL with increasing
  number of threads}\\%[-1mm]
% \multicolumn{6}{c}{{\footnotesize
%   $1/4 (T_1/T_2 + T_2/T_4 + T_4/T_6 + T_6/T_8)$,
%  where $T_i$ is the reasoning time with $i$ threads allocated}}\\
\hline
Finance&3.06&1.88&2.29&2.95&2.73\\
Molecule&1.04&1.24&1.13&1.07&1.05\\
GrossAnatomy&1.41&1.69&1.33&1.32&1.22\\
Skeleton&3.28&1.84&1.65&1.74&1.96\\
ChemistryPrimitive&\textbf{1.46}&\textbf{1.43}&\textbf{1.44}&\textbf{1.51}&\textbf{1.65}\\
FacebookOnto&\textbf{1.81}&\textbf{1.83}&\textbf{1.82}&\textbf{1.61}&\textbf{1.78}\\
Mahabharata&\textbf{1.41}&\textbf{1.46}&\textbf{1.51}&\textbf{1.47}&\textbf{1.56}\\
Transportation&\textbf{1.98}&\textbf{1.57}&\textbf{1.61}&\textbf{1.55}&\textbf{1.59}\\
\hline
\multicolumn{6}{c}{average speedup of RDFox with increasing
  number of threads}\\%[-1mm]
% \multicolumn{6}{c}{{\footnotesize
%   $1/4 (T_1/T_2 + T_2/T_4 + T_4/T_6 + T_6/T_8)$,
%  where $T_i$ is the reasoning time with $i$ threads allocated}}\\
\hline
Finance&1.01&1.04&1.13&1.11&1.19\\
Molecule&0.90&1.03&0.95&1.01&0.99\\
GrossAnatomy&0.87&0.99&0.94&0.96&0.98\\
Skeleton&1.13&1.01&1.06&1.08&1.08\\
ChemistryPrimitive&1.28&1.16&1.14&1.11&1.15\\
FacebookOnto&1.22&1.08&1.07&1.09&1.09\\
Mahabharata&1.24&1.07&1.13&1.10&1.12\\
Transportation&1.14&1.04&1.03&1.08&1.11\\
\hline
\end{tabular}
% \begin{tabular}{crrrrr}
% \hline
% series & 1 & 2 & 3 & 4 & 5\\
% \hline
% &0.54&1.36&0.94&0.99&1.70\\
% Finance&3.06&1.88&2.29&2.95&2.73\\
% &1.01&1.04&1.13&1.11&1.19\\
% \hline
% &0.94&0.79&0.96&0.98&1.07\\
% Molecule&1.04&1.24&1.13&1.07&1.05\\
% &0.90&1.03&0.95&1.01&0.99\\
% \hline
% &0.65&0.48&0.84&0.77&0.83\\
% GrossAnatomy&1.41&1.69&1.33&1.32&1.22\\
% &0.87&0.99&0.94&0.96&0.98\\
% \hline
% &1.11&\textbf{3.08}&\textbf{3.03}&\textbf{3.04}&\textbf{2.75}\\
% Skeleton&3.28&1.84&1.65&1.74&1.96\\
% &1.13&1.01&1.06&1.08&1.08\\
% \hline
% &1.29&1.54&1.68&1.14&0.93\\
% ChemistryPrimitive&\textbf{1.46}&\textbf{1.43}&\textbf{1.44}&\textbf{1.51}&\textbf{1.65}\\
% &1.28&1.16&1.14&1.11&1.15\\
% \hline
% &\textbf{0.23}&\textbf{0.22}&\textbf{0.19}&\textbf{0.24}&\textbf{0.22}\\
% FacebookOnto&\textbf{1.81}&\textbf{1.83}&\textbf{1.82}&\textbf{1.61}&\textbf{1.78}\\
% &1.22&1.08&1.07&1.09&1.09\\
% \hline
% &1.19&1.19&0.99&0.84&0.89\\
% Mahabharata&\textbf{1.41}&\textbf{1.46}&\textbf{1.51}&\textbf{1.47}&\textbf{1.56}\\
% &1.24&1.07&1.13&1.10&1.12\\
% \hline
% &0.24&0.52&0.51&0.65&0.49\\
% Transportation&\textbf{1.98}&\textbf{1.57}&\textbf{1.61}&\textbf{1.55}&\textbf{1.59}\\
% &1.14&1.04&1.03&1.08&1.11\\
% \hline
% \end{tabular}
}
\label{tab:expresult}
\end{table}

\textbf{The Experimental Results}.
We evaluated ParallelDHL and RDFox over the above eight ontology series.
The running environment is a DELL server with a
memory of 16 GiB and 4 cores.
For fairness, we set the same number of threads (i.e., 1,2,4,6 and 8 threads, respectively)
for ParallelDHL and RDFox in each experiment. The resulting reasoning
times\footnote{The experimental results can be found at https://github.com/quanzz/PT.} are
presented in the 16 line graphs (denoted by $lg1$,...,$lg16$) of Figure~\ref{fig:linegraph},
where the abscissa of each line graph records the numbers of ABox assertions,
the ordinate records the reasoning times (in milliseconds),
and each of the five curves in different colors denotes the trend of reasoning time with the
corresponding number of threads allocated (we use line-$k$ to denote the curve
corresponding to $k$ threads).

% We further process the collected data and fill the results in Table~\ref{tab:expresult}, where each cell corresponds
% to a test ontology (see the row label) and a newly generated ABox (distinguished by the column labels);
% the three values from above to below in each cell are: (1) \emph{the minimal time ratio} - the minimal reasoning time
% of ParallelDHL divided by the minimal reasoning time of RDFox;
% (2) \emph{the average speedup of ParallelDHL};\footnote{Suppose $T_i$ is the reasoning time with $i$ threads allocated,
% the average speedup is $\frac{1}{4}(\frac{T_1}{T_2}+\frac{T_2}{T_4}+\frac{T_4}{T_6}+\frac{T_6}{T_8})$.}
% (3) \emph{the average speedup of RDFox}.
% The minimal time ratio describes the performance of ParallelDHL by using RDFox as the baseline.
% The indicator speedup and its derived indicators are the most common tools used for measuring the capacity of parallelism \cite{MotikNPHO14,KazakovKS14,UrbaniKMHB12}.
% Here, we use the average speedup \cite{ichiyoshiK92} to describe the average capacity of parallelism with different
% threads allocated.
% In the following, we give the detailed analysis based on the contents in Figure~\ref{fig:linegraph} and Table~\ref{tab:expresult}.

We further process the collected data and fill the results in Table~\ref{tab:expresult}, where each cell corresponds
to a test ontology (see the row label) and a newly generated ABox
(distinguished by the series column labels). 
The minimal time ratio describes the performance of ParallelDHL by using RDFox as the baseline.
The indicator speedup and its derived indicators are the most common tools used for measuring the capacity of parallelism \cite{MotikNPHO14,KazakovKS14,UrbaniKMHB12}.
Here, we use the average speedup \cite{ichiyoshiK92} to describe the average capacity of parallelism with different
threads allocated.
In the following, we give a detailed analysis based on the contents in Figure~\ref{fig:linegraph} and Table~\ref{tab:expresult}.

\textbf{Analysis of the Experimental Results}.
According to the theoretical results in Section~\ref{sec:ptonto},
materialization of the
ontologies not belonging to $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$
(Finance, Molecule, GrossAnatomy and Skeleton)
may not be tractable in parallel. In other words,
parallel techniques may not work for improving the efficiency of materialization.
This is shown in the line graphs of Figure~\ref{fig:linegraph}.
We can see that, in the line graphs ($lg1$-$lg8$) of the ontology series
that belong to $\mathcal{D}^-_{\textit{\text{dhl}}(\circ)}$, the four lines, line-2, line-4, line-6 and line-8,
intersect to some degree. This situation in the line graphs for Finance and GrossAnatomy
series is more obvious. For example, in line graph $lg1$,
line-2 stays higher than line-6, but lower than line-4.
The intersection of lines indicates that reasoning time cannot be obviously reduced with more threads allocated.
This is also supported by the results of the average speedups.
From Table~\ref{tab:expresult}, we can see that the average speedups of ParallelDHL and RDFox
for ChemistryPrimitive, FacebookOnto, Mahabharata and Transportation series
are more stable compared with the other ones, i.e., they are averagely 1.4
(it means that the reasoning time can be reduced averagely by 1.4 times with two more threads allocated)
for ParallelDHL, and, 1.1 for RDFox. For the four ontology series that belong to $\mathcal{D}^-_{\textit{\text{dhl}}(\circ)}$,
the average speedups have a higher volatility. For example,
Finance and Skeleton series lead to high average speedups that reach up to 3,
while, for Molecule and GrossAnatomy series, the average speedups are even lower than 1.
In summary, from the experiments on the test ontology series,
parallelism leads to a more effective improvement for materializing the ontology series
that belong to $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$ compared with the ones in $\mathcal{D}^-_{\textit{\text{dhl}}(\circ)}$.

By analyzing the experimental results of ParallelDHL and RDFox, we have that
ParallelDHL is a competitive system.
From Table~\ref{tab:expresult}, we can find that most of the minimal time ratios are close to 1.
This means that the reasoning times of ParallelDHL are close to that of RDFox in parallel.
In particular, for GrossAnatomy, FacebookOnto and Transportation series, ParallelDHL performs
much better than RDFox; therein, for materializing FacebookOnto-3, ParallelDHL
only costs one fifth of the minimal reasoning time of RDFox.
On the other hand, for most ontology series,
the average speedups of ParallelDHL are overall higher than that of RDFox.
The main difference of ParallelDHL and RDFox lies in that ParallelDHL applies
the optimizations designed for Algorithm~$\mathsf{A}_{\text{opt}}$ (see Section~\ref{sec:ptclass}).
The higher average speedups of PrallelDHL also verifies the validity of the optimizations
used in Algorithm~$\mathsf{A}_{\text{opt}}$.
One may note that, ParallelDHL is averagely 3 times slower than RDFox when handling Skeleton series.
The main reason is that the computation of \texttt{rch} relations
occupies a large amount of time. We have checked that, the situation of path twisting occurs in Skeleton series.
This makes the the optimization based on \texttt{rch} relations invalid as discussed
in Section~\ref{sec:ptonto}.

### 2018-02-01 parallel-tractability-J-v9/parallelTractabilityOntMat.tex
worked on experimental section, fixed table layout and captions

In this section, we study the issue of parallel tractability
for materialization of DL-Lite and DHL (DHL($\circ$)) ontologies
based on Algorithm~$\mathsf{A}_{\text{opt}}$. 1We show that, for any DL-Lite$_{\text{core}}$ or DL-Lite$_\mathcal{R}$ ontology $\mathcal{O}$,
there exists a poly-logarithmically bounded function $\psi$
such that Algorithm~$\mathsf{A}_{\text{opt}}^{\psi}$ can handle materialization of $\mathcal{O}$.
However, for DHL and DHL($\circ$), there exist ontology classes such that Algorithm~$\mathsf{A}_{\text{opt}}^\psi$
does not work. We illustrate the reason why Algorithm~$\mathsf{A}_{\text{opt}}^\psi$ cannot always
work by studying specific cases.
Further, we propose to restrict the usage of DHL and DHL($\circ$) in order to achieve parallel tractability
of materialization.

@@

\begin{table}
\begin{center}
\begin{tabular}{lcc}
\multicolumn{3}{c}{\textbf{Table~\ref{tab:rdfs}: The RDFS statements and the corresponding DL axioms}}\\
\hline
&~~~~~~~~~~~~~RDFS statements~~~~~~~~~~~~~&~~~~~~~~~~~~~~Axiom~~~~~~~~~~~~~~\\
\hline
\hline
(1)&$P~~\texttt{rdfs:domain}~~C$& $\top\sqsubseteq\forall P^-.C$\\

@@

(6)&$a~~R~~b$& $R(a,b)$\\
\hline
\end{tabular}
\label{tab:rdfs}
\end{center}
\end{table}

>>>>

In this section, we study the issue of parallel tractability
for materialization of DL-Lite and DHL (DHL($\circ$)) ontologies
based on Algorithm~$\mathsf{A}_{\text{opt}}$. We show that, for any DL-Lite$_{\text{core}}$ or DL-Lite$_\mathcal{R}$ ontology $\mathcal{O}$,
there exists a poly-logarithmically bounded function $\psi$
such that Algorithm~$\mathsf{A}_{\text{opt}}^{\psi}$ can handle materialization of $\mathcal{O}$.
However, for DHL and DHL($\circ$), there exist ontology classes such that Algorithm~$\mathsf{A}_{\text{opt}}^\psi$
does not work. We illustrate the reason why Algorithm~$\mathsf{A}_{\text{opt}}^\psi$ cannot always
work by studying specific cases.
Further, we propose to restrict the usage of DHL and DHL($\circ$) in order to achieve parallel tractability
of materialization.

@@

\begin{table}
\centering
\caption{RDFS statements and the corresponding DL axioms}
{\setlength{\tabcolsep}{6mm}
\begin{tabular}{lcc}
\hline
& RDFS statements & Axiom\\
\hline
\hline
(1)&$P~~\texttt{rdfs:domain}~~C$& $\top\sqsubseteq\forall P^-.C$\\

@@

(6)&$a~~R~~b$& $R(a,b)$\\
\hline
\end{tabular}}
\label{tab:rdfs}
\end{table}
